Title: Fish do not need Visas: the transboundary nature of world fished species (70/96 character maximum )

### 2019-11-08 Manuscript_Draft.Rmd
Drafted some more results

We combined a series of datasets and species distribution models to estimate the number of transboundary fished species and their contribution to food security and economics of the world. We adopted the UN definition of transboundary stocks (shared by neighboring EEZs) but carried the analysis at the species level, rather than stocks (populations within a species), due to lack of stock-specific spatial and ecological information on all of the fished species. Our results suggest that there are XXX transboundary species in the world that account for XXX t landings generating USD xxx in fishing revenue to a total of USD xxxx in the fishing sector. These numbers represent an important contribution to global fish production and revenue estimated in XXX and XXX, respectively, highlighting the importance of transboundary stocks worldwide[@REf][@REf].

Regional variation show higher frequency of transboundary species within EEZs along the South China sea, western central Atlantic, and the Mediterranean sea, followed by the Eastern north Atlantic (Fig 1). The overall fish production of transboundary species is relative lower than the number of species within each EEZ, with the exception of some punctual cases like Peru, Chile, Norway, and Ireland where fewer species produce higher landings (Fig 1).

@@

It is not surprising then, that countries within this regions hold the most amount of transboundary species. Some African, Asian, and European countries house over 100 species within their EEZs. Only the US presents over 100 transboundary species outside those continents, once all three sub-EEZs (Atlantic and Pacific coast as well as Alaska arctic and subarctic) are accounted for (Fig. 2).

@@

<!-- Will run statistics to see if I can show the "why" of trans spp -->
There is no clear visual pattern regarding transboundary and discrete species, however, from the three countries that house the most amount of discrete stocks, Australia, New Zeland and the US are among the larger in terms of EEZ length, in the other hand, Brazil has only two neighbors along its large latitudinal coast, suggesting not a lot of connection points for sharing species.
<!-- End of statistics paragraph -->

>>>>

We combined a series of datasets and species distribution models to estimate the number of transboundary fished species and their contribution to food security and economics of the world. We adopted the UN definition of transboundary stocks (shared by neighboring EEZs) but carried the analysis at the species level, rather than stocks (populations within a species), due to lack of stock-specific spatial and ecological information on all of the fished species. Our results suggest that there are `r Total_trans_n` transboundary species in the world, wich is **something** related to what Caddy (1997) estimated and more than double of that of Teh and Sumaila (2010) [@Teh].

Tranboundary fisheries are quite important worldwide in terms of both landings and revenue as they are responsable for XXX tones of landings generating USD `r Total_Trans_Rev` million in fishing revenue. This represents XX% of the catch and `r round(Per_Trans_Rev,2)`% of total revenue generated by those fisheries in all EEZs of the world. All these values are higher than previouslly reported by Teh and Sumaila (2010) up to 20006 where the maximum amount was 45% in the mid-1990s [@Teh]. The importance of these species exceeds in EEZs like the United States, Peru and Norway where fewer species are responsable for higer landing values (Fig 1).

@@

Transboundary fisheries are vast in regions whith historic fisheries-related conflicts [@Spikjers]. Regional variation show higher frequency of transboundary species within EEZs along the South China sea, western central Atlantic, and the Mediterranean sea, followed by the Eastern north Atlantic (Fig 1). Some of these regions have been identified as most conflictive since 1970s with conflicts arrousing to even armed conflicts. Moreover, as climate change shifts species distributions and new transboundary species arrose [@Pinsky] as seen in past cases like the Pacific salmon [@REF], or the Macrrell waars [@Spijkers]. The joint management of shared stocks could be key to not only sustain fisheries into the future but also avoid international conflict [@REF].

Management of shared stocks have to be something. Although catch trends do not reflect stock status [@REf_Haters], they can inform provide a panoram of the situation when no other data is available [@Pauly]. In the case of transboundary species, we did not find a strong pattern suggesting a specific case (Fig 2).

@@
